Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector

### 2018-08-03 documentation/paper.tex
Added results for raw waveform CNN

Supervised learning was used to directly replace Acoustic Parameter. It takes advantage of labeled sets of neutron and alpha events, which are used as training. Several different neural network configurations were applied to this task.

@@

A convolutional neural network was used for analyzing the raw waveform directly, without any preprocessing whatsoever. This avoids any destruction of information, and should be possible, in theory, for a neural network to handle. For this task, a very deep 1-dimensional fully convolutional neural network was applied. The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms.

Two different configurations were tested: one that operates strictly on the waveform with no position corrections of any kind (furthermore $DeepConv(\omega)$) and one with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$).

>>>>

Supervised learning was the first step in the optimization procedure. It makes use of fully labeled calibration sets, which are used for training. Performance of these supervised neural networks is judged based on its accuracy at replication of the provided labels on a validation set (which is set aside and not trained on).

@@

A convolutional neural network was used for analyzing the raw waveform directly, without any preprocessing whatsoever. This avoids any destruction of information, and should be possible, in theory, for a neural network to handle. For this task, a very deep 1-dimensional fully convolutional neural network was applied. The architecture was inspired by the M34-res network \cite{verydeepconvnets} for analysis of raw waveforms. (Its hyperparameters were modified significantly through manual empirical testing as well as grid searches.)

Two different configurations were tested: one that operates strictly on the waveform with no position corrections of any kind (furthermore $DeepConv(\omega)$) and one with an additional position input on a lower layer such that it should be able to learn to incorporate position corrections into its outputs ($DeepConvWithPos(\omega, \chi)$). The second configuration was introduced because the position corrections applied to $\beta_{8}$ in previous PICO-60 research were applied only to broad frequency bands and are thus not practical to apply to a full resolution waveform.

During early tests with variations on $DeepConv(\omega)$, it was revealed to have an incredible capability to separate run types based on unwanted biases. It was able to separate the AmBe calibration run from combined background runs with 96\% accuracy, and to separate all calibration runs from background runs with 91\% accuracy. However, during these tests, the white noise at the beginning and hydraulic recompression sounds at the end were not cropped off, and the network was shown to be overfitting on this information.

Once the audio was cropped properly, $DeepConv(\omega)$ achieved a maximum of 77\% accuracy on validation data. Using the same network architecture (except for the input layer), $DeepConvWithPos(\omega, \chi)$ managed 85\%. The maximum obtained on a grid search with position correction was 90\%. While this is not a drastic improvement, it is a fairly strong indication that the network is able to extract meaningful information from the position, which may be used in an internal position correction of the volume. That said, neither of these accuracy statistics are particularly good, which does not bode well for this input format as a whole.
